---
title: "Analysis of 100-QTL simulated datasets to test for polygenic selection"
output: html_document
author: Timothy Beissinger
date: 10-6-2017
---

### Summary
This script will analyze drift simulations with a particular value for Meff. It will calculate the false positive rate and the mean R2 corresponding to that value.

### Set preliminary values and load function.
These values will store results of the loop

```{r}
Pvals.drift <- c()
R2 <- c()
source("Ghat.R")

Meff <- 40000 #corresponding to decay bin 0 - 0.05

```

### Start the loop over drift simulations.

Simulations were performed on the IRCF Biocluster using the software QMSim. Each iteration of the loop will load data, manipulate data, conduct a test for selection, and record the p-value.

```{r}
for(sim in 1:100){
	iter <- sprintf("%03i",sim)
	directory="/home/beissingert/data/PolySelect_bigsims/QTL_100_big/r_qtl_100"

	#load genotypes and phenotypes
      	pheno <- read.table(paste(directory, "/DriftPop_data_",iter,".txt",sep=""),header=T,stringsAsFactors=F)
	map <- read.table(paste(directory, "/lm_mrk_",iter,".txt",sep=""),header=T,stringsAsFactors=F)
	geno <- read.table(paste(directory, "/DriftPop_mrk_",iter,".txt-head",sep=""),header=F,stringsAsFactors=F,skip=1,sep="", colClasses=c("numeric","character")) # read genos as characters
	qtlMap <- read.table(paste(directory, "/lm_qtl_",iter,".txt",sep=""),header=T,stringsAsFactors=F)

	#load allele frequencies
      	freqs0<-read.table(paste(directory, "/DriftPop_freq_mrk_",iter,".txt",sep=""),header=T,nrows=nrow(map),fill=T,stringsAsFactors=F)
	freqs20<-read.table(paste(directory, "/DriftPop_freq_mrk_",iter,".txt",sep=""),header=F,skip={4*nrow(map)+1},fill=T,stringsAsFactors=F)

	### Manipulate genotypes by coding to -1,0,1 and so that markers are columns, individuals are rows.
	gen <- matrix(NA,nrow=nrow(map),ncol=nrow(geno))
	gen <- as.data.frame(gen)
	names(gen) <- geno[,1]
	for(i in 1:1000){
	  #print(i)
	  tmp <- as.numeric(unlist(strsplit(geno[i,2],split="")))
	  tmp[which(tmp == 0)] <- -1
	  tmp[which(tmp == 3 | tmp ==4)] <- 0
	  tmp[which(tmp==2)] <- 1
	  gen[,i] <- tmp
	 }
	gen<-t(gen)
	gc()

	### Calculate allele frequencies
	#Gen 0
	names(freqs0)[4]<- "Allele1"
	names(freqs0)[5]<- "Allele2"
	freqs0$Allele2[which(substr(freqs0$Allele1,1,1)==2)] <- "2:1.00000" # put this in the spot for the second allele
	freqs0$Allele1[which(substr(freqs0$Allele1,1,1)==2)] <- "1:0.00000" 
	freqs0$Allele2[which(substr(freqs0$Allele1,3,3)==1)] <- "2:0.00000" ##
	freqs0$Allele1 <- as.numeric(substr(freqs0$Allele1,3,1000))
	freqs0$Allele2 <- as.numeric(substr(freqs0$Allele2,3,1000))

	#Gen 20
	names(freqs20)[4]<- "Allele1"
	names(freqs20)[5]<- "Allele2"
	freqs20$Allele2[which(substr(freqs20$Allele1,1,1)==2)] <- "2:1.00000" # put this in the spot for the second allele
	freqs20$Allele1[which(substr(freqs20$Allele1,1,1)==2)] <- "1:0.00000" 
	freqs20$Allele2[which(substr(freqs20$Allele1,3,3)==1)] <- "2:0.00000"  ##
	freqs20$Allele1 <- as.numeric(substr(freqs20$Allele1,3,1000))
	freqs20$Allele2 <- as.numeric(substr(freqs20$Allele2,3,1000))

	#Calculate change
	change2<-freqs20$Allele2-freqs0$Allele2

	### Setup for function
	geno <- gen
	phen <- as.matrix(pheno[1:1000,10])
	rownames(phen)<-pheno[1:1000,1]
	change=change2
		
	# Run function
	test<- Ghat_func(geno=geno,phen=phen,change=change2,method = "scale", num_eff = Meff,  perms=1000, plot="Both")
	Pvals.drift[as.numeric(iter)] <- test$p.val

	# Record LD value
	R2[sim] <- as.numeric(substr(readLines(paste("../../QTL_100_big/r_qtl_100/DriftPop_ld_decay_",iter,".txt",sep=""))[44],10,15))
}
```

### Report the results

Below are some summary statistics about the results. Note also that I save the workspace.

```{r}
hist(Pvals.drift)
print(Pvals.drift)
print(mean(R2))

length(which(Pvals.drift<=0.05))/length(Pvals.drift)

save.image("QTL-100_L44.RData")
```