---
title: "Analysis of the effect that number of generations has on Ghat with 1000 simulated QTL"
output: html_document
author: Timothy Beissinger
date: 2-15-2018
---

### Summary
We simulated a trait controlled by 1,000 QTL 100 times. For each, one population was selected for an increase in the trait and one population was not selected at all (drift only). Selection or drift persisted for a variable numebr of generations. Herein, we use these simulations to test the power of our test statistic to identify selection on the trait.

### Set preliminary values and load function.
These values will store results of the loop

```{r}
Pvals.10 <- c()
source("Ghat.R")
```

### Number of generations: 10

Simulations were performed on the IRCF Biocluster using the software QMSim. Each iteration of the loop will load data, manipulate data, conduct a test for selection, and record the p-value.

```{r}
for(sim in 1:100){
	iter <- sprintf("%03i",sim)
	directory="/home/beissingert/data/PolySelect_bigsims/Generations/r_gens"

	#load genotypes and phenotypes
      	pheno <- read.table(paste(directory, "/10Gen_data_",iter,".txt",sep=""),header=T,stringsAsFactors=F)
	map <- read.table(paste(directory, "/lm_mrk_",iter,".txt",sep=""),header=T,stringsAsFactors=F)
	geno <- read.table(paste(directory, "/10Gen_mrk_",iter,".txt-head",sep=""),header=F,stringsAsFactors=F,skip=1,sep="", colClasses=c("numeric","character")) # read genos as characters #
	qtlMap <- read.table(paste(directory, "/lm_qtl_",iter,".txt",sep=""),header=T,stringsAsFactors=F)

	#load allele frequencies
      	freqs0<-read.table(paste(directory, "/10Gen_freq_mrk_",iter,".txt",sep=""),header=T,nrows=nrow(map),fill=T,stringsAsFactors=F)
	freqs20<-read.table(paste(directory, "/10Gen_freq_mrk_",iter,".txt",sep=""),header=F,skip={nrow(map)+1},fill=T,stringsAsFactors=F) #

	### Manipulate genotypes by coding to -1,0,1 and so that markers are columns, individuals are rows.
	gen <- matrix(NA,nrow=nrow(map),ncol=nrow(geno))
	gen <- as.data.frame(gen)
	names(gen) <- geno[,1]
	for(i in 1:1000){
	  #print(i)
	  tmp <- as.numeric(unlist(strsplit(geno[i,2],split="")))
	  tmp[which(tmp == 0)] <- -1
	  tmp[which(tmp == 3 | tmp ==4)] <- 0
	  tmp[which(tmp==2)] <- 1
	  gen[,i] <- tmp
	 }
	gen<-t(gen)
	gc()

	### Calculate allele frequencies
	#Gen 0
	names(freqs0)[4]<- "Allele1"
	names(freqs0)[5]<- "Allele2"
	freqs0$Allele2[which(substr(freqs0$Allele1,1,1)==2)] <- "2:1.00000" # put this in the spot for the second allele
	freqs0$Allele1[which(substr(freqs0$Allele1,1,1)==2)] <- "1:0.00000" 
	freqs0$Allele2[which(substr(freqs0$Allele1,3,3)==1)] <- "2:0.00000" ##
	freqs0$Allele1 <- as.numeric(substr(freqs0$Allele1,3,1000))
	freqs0$Allele2 <- as.numeric(substr(freqs0$Allele2,3,1000))

	#Gen 20
	names(freqs20)[4]<- "Allele1"
	names(freqs20)[5]<- "Allele2"
	freqs20$Allele2[which(substr(freqs20$Allele1,1,1)==2)] <- "2:1.00000" # put this in the spot for the second allele
	freqs20$Allele1[which(substr(freqs20$Allele1,1,1)==2)] <- "1:0.00000" 
	freqs20$Allele2[which(substr(freqs20$Allele1,3,3)==1)] <- "2:0.00000"  ##
	freqs20$Allele1 <- as.numeric(substr(freqs20$Allele1,3,1000))
	freqs20$Allele2 <- as.numeric(substr(freqs20$Allele2,3,1000))

	#Calculate change
	change2<-freqs20$Allele2-freqs0$Allele2

### Setup for function
    	geno <- gen
        phen <- as.matrix(pheno[1:1000,10])
        rownames(phen)<-pheno[1:1000,1]
	change=change2
	perms <- 10000
	blockSize=100000/100

### Determine num_eff
    	potentials <- c(40000,13333,6667,4000,2857,2222,1818,1538,1333,1176,1053,667,400,286,222)
    	ldTab <- read.table(paste(directory,"/10Gen_ld_decay_",iter,".txt",sep=""),skip=43,nrows=15)
	ld <- as.numeric(substr(ldTab[,2],1,6))
	Meff <- potentials[which(ld <= 0.03)[1]]
	if(is.na(Meff)) Meff <- potentials[15]  #
	print(Meff)
	

	# Run function
	test<- Ghat_func(geno=geno,phen=phen,change=change2,method = "scale", num_eff = Meff,  perms=1000,plot="Both")
	Pvals.10[as.numeric(iter)] <- test$p.val
}
```






### Report the results

Below are some summary statistics about the results. Note also that I save the workspace.

```{r}

hist(Pvals.10)
print(Pvals.10)
length(which(Pvals.10<=0.05))/length(Pvals.10)


save.image("Gens10.RData")
```